---
title: "Numbats in Australia"
author: "Cindy Fajardo"
date: "`r Sys.Date()`"
output: rmdformats::readthedown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../Output/")
```
# Introduction

Welcome to this weeks Tidy Tuesday. This week we will be working with data that was collected on Australia about numbats. They are an endangered species and people are obsessed with them, because they are cute. I am attempting to answer the following questions:

-  Where do you find numbats in Australia?
-  Are they more frequent in the summer or winter?

Wish me luck. Here is a cute numbat pic for inspo. Toodles!
![Numbat Pic](http://t3.gstatic.com/licensed-image?q=tbn:ANd9GcTK2w-LXguKXsOvOyWSDZP96EWf2JyeisaZ_Ks3EM08Kl9UFm6C4_utWfScEAofTlz_TQLAGq1OoTHWgRE)

# Libraries
```{r Libraries}
library(ggmap)
library(tidyverse)
library(here)
library(ggsn)
library(emojifont)
library(readr)
library(nationalparkcolors)
library(praise)
library(devtools)
library(beyonce)
library(ggthemes)
library(hrbrthemes)
library(ggridges)
library(nationalparkcolors)
library(tidytuesdayR)
library(ggplot2)
library(viridis)
library(lubridate)
```
# Data
You can manually read in the data using the code
```{r}
numbats <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-03-07/numbats.csv')
```
# Question 1: Where do you find numbats in Australia?
I used the code below to use google maps to provide the map of Australia so I can then place the locations of where there were many Numbat sightings. We can see from the map that there are many sightings in southwest Australia.
```{r Numbat Map}
Australia <- get_map("Australia", zoom = 4, maptype = "terrain") #satellite map of australia
ggmap(Australia) + #makes map
  geom_point(data = numbats, #uses data for points
             aes(x = decimalLongitude, #uses longitude from numbat data
                 y = decimalLatitude)) #uses latitude from numbat data

```

# Question 2
I used the raw data to place sightings on the maps. Now lets clean up the data to answer the second question.
```{r Sightings in Different Seasons}
clean_numbats <- numbats %>% #numbats data
  drop_na(month) %>% #drop na's
  select(decimalLatitude, decimalLongitude, month, eventDate) %>% #select columns I want to use
  mutate(Month = month(eventDate), #extract month out of UTC column
         season  = case_when( # make a season
           Month %in% c(12,1,2) ~ "winter",
           Month %in% c(3,4,5) ~ "spring",
           Month %in% c(6,7,8) ~ "summer",
           Month %in% c(9,10,11) ~ "fall"),
         season = factor(season), #make season a factor
         season = fct_relevel(season, c("winter","spring","summer","fall"))) #arrange as seasons fall through the year

ggplot(clean_numbats,
       aes(x = Month,
           y = season,
           fill = season))+
  geom_density_ridges(alpha = 0.6)+ #ridges plot
  theme_ridges()+
  theme(legend.position = "none")+ #deletes legend 
  labs(y = "Season",
       title = "Density of Numbat Sightings per Season")
```

