---
title: "Egg_Production_Tidy_Tuesday"
author: "Cindy Fajardo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../Output/")
```

# Introduction
This weeks TidyTuesday will be on a data set that tracks cage-free egg production in the US from July 2016 through February 2021. The data was collected by the United States Department of Agriculture (USDA). This data is only a subset of the entire data that has been collected by USDA which spans from December 2007 to present time.

# Libraries
```{r, Libraries}
library(tidyverse)
library(here)
library(NatParksPalettes)
library(devtools)
library(beyonce)
library(ggthemes)
library(hrbrthemes)
library(viridis)
```


# Read in Data
```{r, Read in Data}
eggproduction  <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-04-11/egg-production.csv') # used this one
cagefreepercentages <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-04-11/cage-free-percentages.csv')

```

# Data Wrangling
```{r, Data Wrangling}
eggproduction_clean <- eggproduction %>% 
  select(observed_month, prod_type, prod_process, n_eggs) %>% # selected columns i will work with
  mutate(year = year(observed_month), # separated year
    month = month(observed_month)) %>% # separated month
  filter(year %in% c(2016, 2017,2018, 2019)) # filtered for years I will be using

eggproduction_clean$egg_type_process<-paste(eggproduction_clean$prod_type, eggproduction_clean$prod_process, sep="_") #pasted production type and production process to create new labels for type of egg

eggproduction_clean$year_month<-paste(eggproduction_clean$year, eggproduction_clean$month, sep="_") ##pasted production type and production process to create new labels for type of egg

```

# Plot
```{r, Plot}
egg_type_labs <- list(
  'hatching eggs_all'="Hatching Eggs",
  'table eggs_all'="All Table Eggs",
  'table eggs_cage-free (non-organic)'="Table Eggs Non-Organic",
  'table eggs_cage-free (organic)'="Table Eggs Organic"
) #created a list with the labels for the egg types

egg_labeller <- function(variable,value){
  return(egg_type_labs[value])
} #created a funtion to create labels


eggproduction_clean %>% 
  group_by(year, egg_type_process) %>% # grouped by year and egg type
  summarise(avg_n_eggs = mean(n_eggs)) %>% #averaeged the number of eggs per year
  ggplot(aes(x = year,
             y= avg_n_eggs,
             fill = egg_type_process))+
  geom_col()+
  facet_wrap(~egg_type_process,
             labeller=egg_labeller)+ #inserted my function here to change facet labels
  theme_bw()+
  guides(fill = FALSE)+ # hide legend
  labs(title = "Average of Type of Egg Produced from 2016-2019",
       x = "Year",
       y = "Average Egg Production")+
  scale_y_continuous(breaks=c(2000000000, 4000000000, 6000000000, 8000000000), labels=c('2 Billion', '4 Billion', '6 Billion', '8 Billion')) # created breaks and labels for y axis
```

